ODGI: understanding pangenome graphs

### 2021-10-21 manuscript/main.tex
File format section

\subsection{Converting to standard text formats}
\label{sec:text}

% pjotr & andrea

% view / paths
Allows us to export basic data of the graph (genomes and graph structure).

\textit{odgi view} can output GFA and internal graph representations \\
\textit{odgi stats} outputs statistics in TSV or YAML. \\
\textit{odgi bin} can output JSON. \\
\textit{odgi degree/position/depth/untangle/tips} can output BED. \\
\textit{odgi untangle} outputs PAF.

Figure: .dot showing the relationship between all the tools and the data formats we use.

>>>>

\subsection{Converting to standard file formats}
\label{sec:text}

% pjotr & andrea
ODGI maintains its own efficient binary format for storing graphs on disk.
ODGI also supports reading and writing the textual GFA file format for interchange with other tools such as with the VG toolkit toolkit~\citep{Garrison:2018}.

\textit{odgi view} can convert a graph in ODGI binary format to the standard GFAv1 format. It can reveal a graph’s internal structures for e.g. debugging processes.

\textit{odgi stats} outputs pangenome statistics in TSV or YAML textual file formats.
Among other metrics, it can calculate the number of nodes, edges, paths and the total nucleotide length of the graph.

\textit{odgi bin} generates a summarized view of gigabase scale graphs providing tab delimited or JSON `bins'. Also it is possible to output a FASTA file.

\textit{odgi degree/position/depth/untangle/tips} all can output the BED format~\citep{Quinlan_2010}.

\textit{odgi untangle} outputs the Pairwise mApping Format (PAF). PAF is a text format describing the approximate mapping positions between two set of sequences.
The odgi untangle command projects paths into a reference-relative BEDPE file, decomposing paralogy relationships.
During this process, it is capable of untangling loopy region resulting in linearized pairs of regions in the BEDPE file. A self dotplot assists in debugging and understanding the untangle process.
